---
title: "Exercise 1"
author: "James Barr"
date: "2/17/2020"
output:
    md_document:
    variant: markdown_github

---

# ABIA Data: Problem 1
## Question: Where did the most flights come from during SXSW & how does that differ from a usual week?
### Reading in the data & necessary packages
```{r Load, echo =FALSE}
library(mosaic)
library(tidyverse)
ABIA = read.csv('ABIA.csv')
```
### Creating New Variables
```{r Variables}
ABIA$Date <- as.Date(with(ABIA, paste(Year, Month, DayofMonth,sep="-")), "%Y-%m-%d")
ABIA$Week = strftime(ABIA$Date, format = "%V")
```
### Filter data down to flights coming to Austin
```{r ToAus}
ToAus = filter(ABIA, Dest == "AUS")
SXSWFlightMonth = filter(ToAus, ToAus$Month == 3)
SXSWFlight = filter(SXSWFlightMonth, between(SXSWFlightMonth$DayofMonth,10,16))
```
### Filter data to just flights to Austin in Spring
```{r ToAusSpring}
Spring_list = c('07','08','09', '10','11','12','13','14','15')
Spring = ToAus %>%
  filter(Week %in% Spring_list) 
SXSWWeek = ToAus %>%
  filter(Week == '11')
```
###How many flights came to Austin each week in Spring
```{r nrowSpring}
nrow(Spring)/9
```
### Create filtered data set for only flights to Austin in SXSW week
```{r SXSWWeekFilter}
SXSWWeek = ToAus %>%
  filter(Week == '11')
nrow(SXSWWeek)
```
### Plot of how many flights come in each week to Austin in Spring
```{r Longitude}
Spring %>%
  mutate(SXSWWeek = ifelse(Week == 11, T,F)) %>%
ggplot(aes(x=Week)) +
  geom_bar(aes(fill = SXSWWeek))+
  labs(x = 'Week', y = 'Number of Flights', title = 'Flights per Week to ABIA in Spring 2008')
```
### Plot of where flights come from throughout spring
```{r SpringOrigin}
SpringOriginBarPlot = Spring %>%
  ggplot(aes(x=Origin)) +
  geom_bar()+
  coord_flip() +
    labs(x = 'Origin', y = 'Number of Flights', title = 'Origins of flights to ABIA in Spring 2008')

SpringOriginBarPlot
```
### Plot of where flights come from throughout SXSW
```{r SXSWOrigin, echo =FALSE}
SXSWOriginBarPlot = ggplot(data = SXSWWeek) +
  geom_bar(aes((x=Origin))) +
    coord_flip() +
      labs(x = 'Origin', y = 'Number of Flights', title = 'Origins of flights to ABIA during SXSW 2008')

SXSWOriginBarPlot
```
##Conclusion
```{r Conclusion}
'There is not a huge difference between SXSW and a normal Spring week in terms of how many flights are coming in and where the flights are coming from.
In both periods, Dallas sent by far the most flights to Austin compared to any other city.'