---
title: "Problem2"
author: "James Barr"
date: "2/17/2020"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,read}
creatinine = read.csv('creatinine.csv')
```
#Question1: What creatinine clearance rate should we expect, on average, for a 55-year-old?
```{r,CreatModel}
lm1 = lm(creatclear ~ age, data=creatinine)
predict(lm1,list(age=55))
```
##Answer: The expected Creatinine clearnace rate for a 55 year old is 113.7228.

#Question2: How does creatinine clearance rate change with age? (This should be a number with units ml/minute per year.)
```{r,CreatCoef}
coef(lm1)
```
##Answer: On average,creatinine clearance rate decreases by 0.6198159ml/minute per year

#Question3: Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112?
````{r, "healthier"}
predict(lm1,list(age=c(40,60)))
#40 year old expected CCR is 123.0203
#60 year old expected CCR is 110.6240

#Formulas Below = Actual - Predicted
135-123.0203
#40 year old CCR is 11.9797mL/min healthier than average
112-110.6240
#60 year old CCR is 1.376mL/min healthier than average
```
###Answer: 40 year old has a healthier CCR for his age than the 60 year old